Detailed disease progression of 97 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis

### 2020-10-16 manuscript/manuscript.Rmd
Further work on risk scores
Improved progression figure

---
title: "Detailed disease progression of 97 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis" 
output: 
  #rticles::arxiv_article
  #pdf_document
  bookdown::pdf_document2: 
     toc: false
  # bookdown::pdf_book: 
  #    base_format: rticles::arxiv_article
abstract: "We collected a multi-centric retrospective dataset of patients who were admitted to N hospitals in Czech Republic and tested positive for SARS-CoV-2. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. To assess association between treatments, markers and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom. Additionally we performed external validation of several proposed prognostic models for Covid-19 severity."
bibliography: Covid19.json
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
---

@@

sum_percent <- function(x) {
  paste0(sum(x, na.rm = TRUE), " (", scales::percent(mean(x, na.rm = TRUE), accuracy = 1), ")")
}

mean_iqr <- function(x) {
  paste0(round(mean(x, na.rm = TRUE)), " (", round(quantile(x, 0.25, na.rm = TRUE)), " - ", round(quantile(x, 0.75, na.rm = TRUE)), ")")  
}

data_table_1 %>% group_by(group) %>% 
  summarise(
    N = n(),
   # `Distinct sites` = length(unique(hospital_id)),
    Male = sum_percent(sex == "M"),
    `Age (mean, IQR)` = mean_iqr(age),
    `Admitted for Covid` = sum_percent(admitted_for_covid),
    `Took Azithromycin` = sum_percent(ever_az),
    `Ischemic Heart Disease` = sum_percent(ischemic_heart_disease),
    `Takes antihypertensives` = sum_percent(has_hypertension_drugs),	
    `Heart Failure` = sum_percent(heart_failure),
    COPD = sum_percent(COPD),
    Asthma = sum_percent(asthma),
    `Other lung disease` = sum_percent(other_lung_disease),
    Diabetes = sum_percent(diabetes),
    `Renal Disease` = sum_percent(renal_disease),
    `Liver Disease` = sum_percent(liver_disease),
    `Smoking` = sum_percent(smoking),
    `BMI (mean, IQR)` = mean_iqr(BMI),
    Deceased = sum_percent(outcome == "Death"),
    Discharged = sum_percent(outcome == "Discharged"),
    .groups = "drop"
  ) %>%
  mutate(group = factor(group, levels= c("All", "HCQ", "No HCQ", "Measured D-dimer","Measured IL-6"), ordered = TRUE)) %>%
  arrange(group) %>%
  as.data.frame() %>%
  column_to_rownames("group") %>%
  t() %>% 
  knitr::kable(caption = "Patient characteristics. Note that the 'Measured D-dimer' and 'Measured IL-6' are neither exclusive nor exhaustive.", booktabs = TRUE)
```

@@

```{r progression, fig.cap="Disease progression for all patients included in the study as determined by breathing support required. AA = Ambient air, Oxygen = Nasal oxygen, NIPPV = Non-invasive positive-pressure ventilation, ECMO = Extra-corporeal membrane oxygenation."}
my_paste <- function(x) {
  paste0(as.integer(x), collapse = "")
}
wide  %>% 
  filter(!is.na(breathing)) %>%
  arrange(hospital_id, patient_id, day) %>%
  mutate(patient_id = factor(patient_id), patient_id = fct_reorder(patient_id, breathing,  .fun = my_paste, .desc = FALSE)) %>%
  ggplot(aes(x = day, y = patient_id, fill = breathing)) + 
  geom_tile(height = 0.7) + 
  scale_fill_viridis_d("", direction = -1) +
  scale_x_continuous("Days since hospitalization") +
  theme(axis.text.y = element_blank()) +
  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank())

@@

>>>>

---
title: "Detailed disease progression of 97 patients hospitalized with Covid-19 in the Czech Republic: An exploratory analysis" 
output: 
  #rticles::arxiv_article
  #pdf_document
  bookdown::pdf_document2:
     toc: false
  # bookdown::word_document2: 
  #     toc: false
  # bookdown::pdf_book: 
  #    base_format: rticles::arxiv_article
abstract: "We collected a multi-centric retrospective dataset of patients who were admitted to N hospitals in Czech Republic and tested positive for SARS-CoV-2. The dataset contains baseline patient characteristics, breathing support required, pharmacological treatment received and multiple markers on daily resolution. To assess association between treatments, markers and patient outcomes we performed multiverse analysis, observing how the conclusions change between defensible choices of statistical model, predictors included in the model and other analytical degrees of freedom. Additionally we performed external validation of several proposed prognostic models for Covid-19 severity."
bibliography: Covid19.json
date: "Version: `r format(Sys.time(), '%Y-%m-%d')`"
---

@@

data_table_1 %>% group_by(group) %>% 
  summarise(
    N = n(),
    `Distinct sites` = length(unique(hospital_id)),
    Male = sum_percent(sex == "M"),
    `Age (mean, IQR)` = mean_iqr(age),
    `Admitted for Covid` = sum_percent(admitted_for_covid),
    `Took Azithromycin` = sum_percent(ever_az),
    `Took Dexamethasone` = sum_percent(ever_dexamethasone),
    `Took Favipiravir` = sum_percent(ever_favipiravir),
    `Took Remdesivir` = sum_percent(ever_remdesivir),
    `Convalescent plasma` = sum_percent(ever_convalescent_plasma),
    `Ischemic Heart Disease` = sum_percent(ischemic_heart_disease),
    `Takes antihypertensives` = sum_percent(has_hypertension_drugs),	
    `Heart Failure` = sum_percent(heart_failure),
    COPD = sum_percent(COPD),
    Asthma = sum_percent(asthma),
    `Other lung disease` = sum_percent(other_lung_disease),
    Diabetes = sum_percent(diabetes),
    `Renal Disease` = sum_percent(renal_disease),
    `Liver Disease` = sum_percent(liver_disease),
    `Smoking` = sum_percent(smoking),
    `BMI (mean, IQR)` = mean_iqr(BMI),
    Deceased = sum_percent(outcome == "Death"),
    Discharged = sum_percent(outcome == "Discharged"),
    .groups = "drop"
  ) %>%
  mutate(group = factor(group, levels= c("All", "HCQ", "No HCQ", "Measured D-dimer","Measured IL-6"), ordered = TRUE)) %>%
  arrange(group) %>%
  as.data.frame() %>%
  column_to_rownames("group") %>%
  t() %>% 
  knitr::kable(caption = "Patient characteristics. Note that the 'Measured D-dimer' and 'Measured IL-6' are neither exclusive nor exhaustive.", booktabs = TRUE)
```

@@

```{r progression, fig.cap="Disease progression for all patients included in the study as determined by breathing support required."}
my_paste <- function(x) {
  paste0(as.integer(x), collapse = "")
}

data_for_fig_progression <- wide  %>% 
  filter(!is.na(breathing_s), day >= 0) %>%
  arrange(hospital_id, patient_id, day) %>%
  mutate(patient_id = factor(patient_id), patient_id = fct_reorder(patient_id, breathing_s,  .fun = my_paste, .desc = FALSE)) %>%
  mutate(first_half = as.integer(patient_id) < length(levels(patient_id)) / 2 + 1, 
         is_outcome =  breathing_s %in% c("Discharged", "Death"),
         breathing_s = fct_recode(breathing_s, `Ambient air` = "AA")
         )

data_for_fig_progression %>% 
  filter(!is_outcome) %>%
  ggplot(aes(x = day, y = patient_id, fill = breathing_s)) + 
  geom_tile(height = 0.7) + 
  geom_point(data = data_for_fig_progression %>% filter(is_outcome) %>% mutate(outcome = factor(breathing_s, ordered = FALSE)),
             inherit.aes = FALSE, aes(x = day, y = patient_id, shape = outcome)) +
  scale_fill_viridis_d("", direction = -1) +
  scale_x_continuous("Days since hospitalization") +
  scale_shape_manual("", values = c(1, 15) ) +
  facet_wrap(~first_half, scales = "free_y") +
  theme(axis.text.y = element_blank(), strip.background = element_blank(), strip.text = element_blank(),
        legend.direction = "horizontal", legend.position = "bottom") +
  theme(axis.title.y = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank())

@@

## Survey of antimicrobial treatment
